---
title: "House_report"
output:
  pdf_document: default
  html_document: default
date: "2023-01-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# House report


## Data

The data for median house price per square feet is extracted from [website](https://www.rate.com/research). The data for number of home sold and median days on the market is captured from the [website](https://www.redfin.com/city/30748/NY/Bedford/housing-market.)


### Included Area

The excel file created for data analytics contains towns that has the following property:

<br>

#### Target area

This is the area we focus on. There are five areas, namely Scarsdale, Purchase, Rye, Bedford, Chappaqua. Scarsdale is the benchmark, which has a high house price. 

(This is the testing dataset. I create a linear model based on other area to predict the potential house price per square feet. So that it can show if the price still has potential to increase)

<br>

#### High median price per square foot

In accordance to [move.org](https://www.move.org/americas-most-expensive-cities-per-square-foot/) webpage, cities like San Francisco, Boston and New York City have the mosty expensive house price per unit. Along with these cities, I also choose Fremont, CA, Oakland, CA, San Jose, CA and Honolulu as focused cities. These cities can give a clear indication of the important factors that should be considered, if the house price is high, or is going to be high in the target city. 

<br>

#### Other chosen city

Seattle and Huston are also two major cities that have a high demand of investment. These cities can also be act as an benchmarks. River Edge is very similar to Chappaqua, so I include it in the data for comparison. 



<br>
<br>


### Variables

#### Table 1: HousePrice

House price per square feet is included in the first table of the data. The table contains all the house price per square feet from 4th Feburary 2022 to 20th January 2023. The mean value for house price per square feet for each month is given in the aggragate columns in the last 12 rows, from BB to BM column in excel.

<br>

#### Table 2: #ofhousehold

Table 2 consists the number of purchased house unit for every month in the last three years. 

<br>

#### Table 3: Day on the market

Table 3 consists the median time for the housing unit being purchased for every month in the last three years.

<br>

#### Table 4: all info

Table 4 consists all the background information of the housing unit. In order to calculate the change in price over seasons so that the impact of the raise in interest rate can be reviewed, I choose to extract the data from Feb. 18th, May 13th, Aug 12th, Nov 18th and the latest available price, which is Jan 20, 2023. Columns C to K in excel are the calculated house values and the change in price is calculated in column H to K so that we can see the potential impact of rise in interest rate after August.

The rating of each city is also included in the table. This consist the overall rating of each city, as well as the rating for Housing, Education,	Cost of living,	Unemployment,	Commute and	Lifestyle. The rating is in a scale of 12. The scale is represented the following: A+: 12 A: 11 A-: 10 B+: 9 B: 8 B-: 7 C+: 6 C: 5 C-: 4 D+: 3 D: 2 D-: 1. 

The description of the rating is evaluated from the following categories: 

* Housing category: considers home inventory availability, median days on market, total new listing and estimated home sales.

* Cost of living: considers median home price and average cost per square foot. 

* Education rating: evaluate the school grade and the percentage of Bachelor degree or greater people in the city.

* Unemployment: considers employment rate only.

* Life style: considers urban life style, including number of restaurant and bars, businesses and net migrations of the city. 

* Commute rating: evaluate the median travel time to work, and percentage of people go to work by bike, public transit or on foot. 



The table also contains the demographical details of each chosen cities, such as housing unit, median income and populations.

The last few columns contains the average school rating of top five schools in the city and the total number of schools in the city. The schools are classified into Elementary school, Middle school and High school. 

<br>


#### Table 5: Population in thousand

This is the population table for each target cities. 

<br>

#### Table 6: Home for sale

This is the table which consists the number of house unit available for every month in the last two years. 

<br>

#### Table 7: Sale to list 

This is the table that store all of the sale price in relative to the list price in percentage. If the stored value is 100, it means that the sale price is exactly the same as the list value. If the value is higher than 100, than the sale price is higher than the list value. 

<br>

#### Table 4: House price in 3 yr

This is another table for the house price in square feet, but with a scale value of 3 years. All values are the median value of each month.

<br>

#### Table 5: Other potential city

It is an incomplete table which may be some potential cities that are worth to invest in. 


<br>
<br>


## Results

```{r,echo=FALSE,message=FALSE}
library("tidyverse")
library("dplyr")
library("readxl")
library(ggplot2)
library(reshape2)
library(ggfortify)
library(wesanderson)
```


```{r,echo=FALSE}
house_price = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 1)
num_sold = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 2)
medium_time = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 3)
all_info = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 4)
pop = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 5)
num_house = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 6)
sale_to_list = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 7)
hp_three_yr = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 8)
```





```{r,include=FALSE}
all_hp = t(all_info[1:16,3:7])
data_df_allhp = melt(all_hp)
data_df_allhp_named = 
  data_df_allhp %>% 
  mutate(Var2 = ifelse(Var2==1,"Chappaqua",ifelse(Var2==2, "Bedford", ifelse(Var2==3, "Rye", ifelse(Var2==4, "Purchase",ifelse(Var2==5,"Scarsdale",ifelse(Var2==6, "Fremont",ifelse(Var2==7, "Oakland", ifelse(Var2==8, "San Jose", ifelse(Var2==9, "Honolulu", ifelse(Var2==10, "San Francisco", ifelse(Var2==11, "Boston", ifelse(Var2==12,"Washington", ifelse(Var2==13, "Los Angeles", ifelse(Var2==14, "Houston",ifelse(Var2==15,"Seattle","River Edge"))))))))))))))),
         ind = rep(1:5,16),
         HousePrice = Var2)
ggplot(data_df_allhp_named, aes(x=ind, y=value, color = HousePrice))+
  geom_line()+
  labs(
    y="house price",
    x="time"
  )+
  scale_x_continuous(
    labels = c("1st_2022","2nd_2022","3rd_2022","4th_2022","1th_2023"))
```


### Clustering results

```{r,warning=FALSE}
selected_info =
  all_info %>% 
  janitor::clean_names() %>% 
  select(overall_rating, housing, education, cost_of_living, unemployment, commute, lifestyle) 
rownames(selected_info) = c("Chappaqua", "Bedford","Rye","Purchase","Scarsdale","Fremont", "Oakland", "San Jose", "Honolulu", "San Francisco", "Boston", "Washington","Los Angeles", "Houston","Seattle","River Edge")

pca_res <- prcomp(selected_info, scale. = TRUE)




df <- as.data.frame(pca_res$x)
df$city <- rownames(df)

ggplot(df, aes(PC1, PC2, color = city)) + 
  geom_point(size = 3) +
  geom_text(aes(label = city) , vjust = 2) +
  lims(x = c(-4, 3), y = c(-3, 3)) +
  theme_bw() +
  theme(legend.position = "none") 
```

Honolulu has a rating that is close to the four target areas, so I include the city in the price time series plot. Seattle is also included for a reference as it has a high house price per square feet and it is a well developed city.

```{r,include=FALSE}
autoplot(pca_res, labels = rownames(selected_info))
```

### House price time series


```{r,warning=FALSE}
time = colnames(house_price)[3:54]
df1 = t(house_price[,3:54])
data_df1 = melt(df1)
data_df1_named = 
  data_df1 %>% 
  mutate(Var2 = ifelse(Var2==1,"Chappaqua",ifelse(Var2==2, "Bedford", ifelse(Var2==3, "Rye", ifelse(Var2==4, "Purchase",ifelse(Var2==5,"Scarsdale",ifelse(Var2==6,"Honolulu",ifelse(Var2==7,"Seattle","River Edge"))))))),
         ind = rep(1:length(time),8),
         HousePrice = Var2)
data_df1_named_change = data_df1_named%>% filter(HousePrice!="Rye" & HousePrice!="Honolulu")
ggplot(data_df1_named_change, aes(x=ind, y=value, color = HousePrice))+
  geom_line(size=1)+
  labs(
    y="house price",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time),5),
    labels = c("Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Nov","Dec","jan"),
    limits = c(1,length(time)))+
  scale_color_brewer(palette="Paired")
```

```{r}
data_df1_named %>% group_by(Var2) %>% summarize(mean_price = mean(value))
```


According to the time series plot for house price per square foot, Bedford has the most fluctuated price over the time. It is one of the most competitive homes sell in 18 days. There is a net population gain in Bedford, with a net search flow of approximately 1100. The house price for Bedford decreases after the increment of interest rate. Rye has the highest house price per square foot among target cities (Rye, Purchase, Chappaqua, Bedford) , and it is relatively stable. The house price per square feet for Purchase does not decrease after the rise in interest rate and it has a persistent increasing trend. 


#### HousePrice 3 year

```{r,warning=FALSE}

timed = colnames(hp_three_yr)[3:38]
timelist = seq(1,length(timed),5)
#timelist = append(timelisttemp,length(timed))
dfnum = t(hp_three_yr[,3:38])
data_dfnum = melt(dfnum)
data_dfnum_named = 
  data_dfnum %>% 
  mutate(Var2 = ifelse(Var2==1,"Chappaqua",ifelse(Var2==2, "Bedford", ifelse(Var2==3, "Rye", ifelse(Var2==4,"Purchase","Scarsdale")))),
         ind = rep(1:length(timed),5),
         hp_threeyr = Var2) %>% 
  filter(hp_threeyr!="Rye" & hp_threeyr!="Seattle")
ggplot(data_dfnum_named, aes(x=ind, y=value, color = hp_threeyr))+
  geom_line(size=1)+
  labs(
    title = "Median house price for three years",
    y="House price three year",
    x="Date"
  )+
  scale_x_continuous(
    breaks = timelist,
    labels = c("Jan_20","Jun_20","Nov_20","Apr_21","Sep_21","Feb_22","Jul_22","Jan_23"),
    limits = c(1,length(timed)))

```





### Number of house


```{r,warning=FALSE}
time2 = colnames(num_sold)[3:37]
num_sold_num = num_sold %>% filter(City_name!="Purchase") %>% filter(City_name!="Seattle") %>% filter(City_name!="Honolulu")
df2 = t(num_sold_num[,3:37])
data_df2 = melt(df2)
data_df2_named = 
  data_df2 %>% 
  mutate(Var2 = ifelse(Var2==1,"Chappaqua",ifelse(Var2==2, "Bedford", ifelse(Var2==3, "Rye", ifelse(Var2==4,"Seattle","River Edge")))),
         ind = rep(1:length(time2),5),
         Numhouse = Var2) %>% 
  filter(Numhouse!="Rye" & Numhouse!="Seattle")
ggplot(data_df2_named, aes(x=ind, y=value, color = Numhouse))+
  geom_line(size=1)+
  labs(
    y="Number of house",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time2),5),
    labels = c("Jan_20","May_20","Oct_20","Mar_21","Aug_21","Jan_22","Jun_22"),
    limits = c(1,length(time2)))

```

```{r}
data_df2_named %>% group_by(Var2) %>% summarize(mean_number = mean(value))
```

Purchase data is not available online. Rye is the city that has the highest number of house available among three target cities. 


### Medium day on the market

```{r,warning=FALSE}
time3 = colnames(medium_time)[3:37]
medium_time_num = medium_time %>% filter(City_name!="Purchase") 
df3 = t(medium_time_num[,3:37])
data_df3 = melt(df3)
data_df3_named = 
  data_df3 %>% 
  mutate(Var2 = ifelse(Var2==1,"Chappaqua",ifelse(Var2==2, "Bedford", ifelse(Var2==3, "Rye",ifelse(Var2==4,"Scarsdale",ifelse(Var2==5,"Honolulu",ifelse(Var2==6,"Seattle","River Edge")))))),
         ind = rep(1:length(time3),7),
         MedianTime = Var2) %>% 
  filter(MedianTime!="Honolulu" & MedianTime!="Seattle" & MedianTime!="Rye")
ggplot(data_df3_named, aes(x=ind, y=value, color = MedianTime))+
  geom_line(size=1)+
  labs(
    y="Medium day on market",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time2),5),
    labels = c("Jan_20","May_20","Oct_20","Mar_21","Aug_21","Jan_22","Jun_22"),
    limits = c(1,length(time2)))+
    scale_color_brewer(palette="Paired")



```

```{r}
data_df3_named %>% group_by(Var2) %>% summarize(mean_day = mean(value))
```



Bedford has a relatively high number of days on the market before sold. 


### Home for sale

```{r}
time4 = colnames(num_house)[3:26]
num_house_num = num_house %>% filter(City_name!="Chappaqua") 
df4 = t(num_house_num[,3:26])
data_df4 = melt(df4)
data_df4_named = 
  data_df4 %>% 
  mutate(Var2 = ifelse(Var2==1, "Bedford", ifelse(Var2==2, "Rye",ifelse(Var2==3,"Purchase",ifelse(Var2==4,"Scarsdale",ifelse(Var2==5,"Honolulu",ifelse(Var2==6,"Seattle","River Edge")))))),
         ind = rep(1:length(time4),7),
         NumHouse = Var2) %>% 
 filter(NumHouse!="Honolulu" & NumHouse!="Seattle" & NumHouse!="Rye"& NumHouse!="Scarsdale")
ggplot(data_df4_named, aes(x=ind, y=value, color = NumHouse))+
  geom_line(size=1)+
  labs(
    y="Number of house on market",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time4),4),
    labels = c("Jan_21","May_21","Sep_21","Jan_22","May_22","Sep_22"),
    limits = c(1,length(time4)))


```
```{r}
data_df4_named %>% group_by(Var2) %>% summarize(mean_number = mean(value))
```





\newpage

## Time series for multiple indicator

#### Note population and population rate

This is the plotting that stores all the house price, number of available house, sold number, number of day and the sale to list price is stored in one plotting for each city. Since they are all in different scale, I transfer them into the same scale using standardization so that all trends are captured in each of the plotting. 

#### Chappaqua




```{r}
std <- function(x){
  return (x-mean(x))/sd(x)
}


#take value from year 2021
cha_num_sold = as.numeric(num_sold[1,][15:37]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
cha_sale_time = as.numeric(medium_time[1,][15:38][-23]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
cha_inventory =  as.numeric(num_house[1,][3:26][-23]) %>% std()
cha_sale_to_list =  as.numeric(sale_to_list[1,][15:37]) %>% std()
cha_hp =  as.numeric(hp_three_yr[1,][15:37]) %>% std()

date = seq(1,23,1)
#time_val = format(date, "%B %Y")
#time_val[23]="December 2022"


chapa_df = as.data.frame(cbind(date,cha_num_sold,cha_sale_time,cha_inventory,cha_sale_to_list,cha_hp)) 
chapa_df_longer <- pivot_longer(chapa_df, c(cha_num_sold,cha_sale_time,cha_inventory,cha_sale_to_list,cha_hp)) %>% na.omit()

ggplot(chapa_df_longer, aes(x=date, y=value, color = name))+
  geom_line(size=1)+
  labs(
    y="Value",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time4),4),
    labels = c("Jan_21","May_21","Sep_21","Jan_22","May_22","Sep_22"),
    limits = c(1,length(time4)))+
  ggtitle("Chappaqua")
```

#### Bedford


```{r,echo=FALSE}
#take value from year 2021
bed_num_sold = as.numeric(num_sold[2,][15:37]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
bed_sale_time = as.numeric(medium_time[2,][15:38][-23]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
bed_inventory =  as.numeric(num_house[2,][3:26][-23]) %>% std()
bed_sale_to_list =  as.numeric(sale_to_list[2,][15:37]) %>% std()
bed_hp =  as.numeric(hp_three_yr[2,][15:37]) %>% std()

date = seq(1,23,1)
#time_val = format(date, "%B %Y")
#time_val[23]="December 2022"


bed_df = as.data.frame(cbind(date,bed_num_sold,bed_sale_time,bed_inventory,bed_sale_to_list,bed_hp)) 
bed_df_longer <- pivot_longer(bed_df, c(bed_num_sold,bed_sale_time,bed_inventory,bed_sale_to_list,bed_hp)) %>% na.omit()

ggplot(bed_df_longer, aes(x=date, y=value, color = name))+
  geom_line(size=1)+
  labs(
    y="Value",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time4),4),
    labels = c("Jan_21","May_21","Sep_21","Jan_22","May_22","Sep_22"),
    limits = c(1,length(time4)))+
  ggtitle("Bedford")
```


#### Rye




```{r,echo=FALSE}
#take value from year 2021
rye_num_sold = as.numeric(num_sold[3,][15:37]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
rye_sale_time = as.numeric(medium_time[3,][15:38][-23]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
rye_inventory =  as.numeric(num_house[3,][3:26][-23]) %>% std()
rye_sale_to_list =  as.numeric(sale_to_list[3,][15:37]) %>% std()
rye_hp =  as.numeric(hp_three_yr[3,][15:37]) %>% std()

date = seq(1,23,1)
#time_val = format(date, "%B %Y")
#time_val[23]="December 2022"


rye_df = as.data.frame(cbind(date,rye_num_sold,rye_sale_time,rye_inventory,rye_sale_to_list,rye_hp)) 
rye_df_longer <- pivot_longer(rye_df, c(rye_num_sold,rye_sale_time,rye_inventory,rye_sale_to_list,rye_hp)) %>% na.omit()

ggplot(rye_df_longer, aes(x=date, y=value, color = name))+
  geom_line(size=1)+
  labs(
    y="Value",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time4),4),
    labels = c("Jan_21","May_21","Sep_21","Jan_22","May_22","Sep_22"),
    limits = c(1,length(time4)))+
  ggtitle("Rye")
```



#### Purchase

```{r,echo=FALSE}
#take value from year 2021
pur_num_sold = as.numeric(num_sold[4,][15:37]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
pur_sale_time = as.numeric(medium_time[4,][15:38][-23]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
pur_inventory =  as.numeric(num_house[4,][3:26][-23]) %>% std()
pur_sale_to_list =  as.numeric(sale_to_list[4,][15:37]) %>% std()
pur_hp =  as.numeric(hp_three_yr[4,][15:37]) %>% std()

date = seq(1,23,1)
#time_val = format(date, "%B %Y")
#time_val[23]="December 2022"


pur_df = as.data.frame(cbind(date,pur_num_sold,pur_sale_time,pur_inventory,pur_sale_to_list,pur_hp)) 
pur_df_longer <- pivot_longer(pur_df, c(pur_num_sold,pur_sale_time,pur_inventory,pur_sale_to_list,pur_hp)) %>% na.omit()

ggplot(pur_df_longer, aes(x=date, y=value, color = name))+
  geom_line(size=1)+
  labs(
    y="Value",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time4),4),
    labels = c("Jan_21","May_21","Sep_21","Jan_22","May_22","Sep_22"),
    limits = c(1,length(time4)))+
  ggtitle("Purchase")
```



#### Scarsdale

```{r,echo=FALSE}
#take value from year 2021
scar_num_sold = as.numeric(num_sold[5,][15:37]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
scar_sale_time = as.numeric(medium_time[5,][15:38][-23]) %>% std()
#get rid of November as not all dataset contains Nov, this is the 25th row
scar_inventory =  as.numeric(num_house[5,][3:26][-23]) %>% std()
scar_sale_to_list =  as.numeric(sale_to_list[5,][15:37]) %>% std()
scar_hp =  as.numeric(hp_three_yr[5,][15:37]) %>% std()

date = seq(1,23,1)
#time_val = format(date, "%B %Y")
#time_val[23]="December 2022"


scar_df = as.data.frame(cbind(date,scar_num_sold,scar_sale_time,scar_inventory,scar_sale_to_list,scar_hp)) 
scar_df_longer <- pivot_longer(scar_df, c(scar_num_sold,scar_sale_time,scar_inventory,scar_sale_to_list,scar_hp)) %>% na.omit()

ggplot(scar_df_longer, aes(x=date, y=value, color = name))+
  geom_line(size=1)+
  labs(
    y="Value",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time4),4),
    labels = c("Jan_21","May_21","Sep_21","Jan_22","May_22","Sep_22"),
    limits = c(1,length(time4)))+
  ggtitle("Scarsdale")
```

\newpage


## Scatter plot for chappaqua and scarsdale

```{r}
all_val = 
  hp_three_yr %>%
  filter(City_name=="Chappaqua" | City_name=="Scarsdale") %>% 
  pivot_longer(
    `43831`:`44927`,
    names_to = "date"
  ) %>% 
  mutate(date_val = as.Date(as.numeric(date),origin="1899-12-30"),
         ind = seq(1,72,1))

ggplot(all_val,aes(x=value, y=ind, color = City_name))+
  geom_point()+
  labs(title="Scatter plot for Chappaqua and Scarsdale")
```


The standard deviation of Chappaqua is ```r round(sd(as.numeric(hp_three_yr[1,][-1:-2])),3)``` with a range of ```r max(as.numeric(hp_three_yr[1,][-1:-2]))-min(as.numeric(hp_three_yr[1,][-1:-2]))``` dollar per square feet. Scarsdale, on the other hand, has a standard deviation of ```r round(sd(as.numeric(hp_three_yr[5,][-1:-2])),3)``` with a range of ```r max(as.numeric(hp_three_yr[5,][-1:-2]))-min(as.numeric(hp_three_yr[5,][-1:-2]))``` dollar per square feet. The table shows all the information:

```{r,echo=FALSE}
all_val %>% 
  group_by(City_name) %>% 
  summarize(
    mean_val = mean(value),
    sd_val = sd(value),
    range_dollar = max(value)-min(value)
  )
```



\newpage



## Mean value for each half year



```{r,echo=FALSE}
summary_half_yr = 
  hp_three_yr %>%
  pivot_longer(col = `43831`:`44927`,names_to = "date") %>%
  mutate(year_half = ifelse(date<44012,"first_2020",ifelse(date<44196,"second_2020",ifelse(date<44377,"first_2021",ifelse(date<44550,"second_2021",ifelse(date<44742,"first_2022","second_2022")))))) %>% 
  group_by(City_name,year_half) %>% 
  summarize(
    mean_val = round(mean(value),2)
  ) %>% 
  ungroup() %>% 
  pivot_wider(
    names_from=year_half,
    values_from = mean_val)
summary_half_yr
```



```{r,echo=FALSE}
cleaned_summ = summary_half_yr %>% 
  pivot_longer(col = first_2020:second_2022, names_to = "year") %>% 
  mutate(time = ifelse(year == "first_2020",1,ifelse(year == "second_2020",2,ifelse(year == "first_2021",3,ifelse(year == "second_2021",4,ifelse(year == "first_2022",5,6))))))

ggplot(cleaned_summ, aes(x=time, y=value, color = City_name))+
  geom_line(size=1)+
  labs(
    y="house price",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,6,1),
    labels = c("2020_1st","2020_2nd","2021_1st","2021_2nd","2022_1st","2022_2nd"),
    limits = c(1,6))+
  scale_color_brewer(palette="Paired")+
  labs(title="Time series plot for mean value every half year")
```


\newpage

## Index

### Yearly index

```{r,echo=FALSE}
yearly_index = 
  summary_half_yr %>% 
  mutate_at(vars(first_2020:second_2022), list(index = ~./first_2020)) %>% 
  select(City_name,ends_with("index")) %>% 
  select(City_name, first_2020_index, second_2020_index,first_2021_index, second_2021_index,first_2022_index, second_2022_index)
  
yearly_index
```

Yearly index plot:

```{r,echo=FALSE}
cleaned_year_index = yearly_index %>% 
  pivot_longer(col = first_2020_index:second_2022_index, names_to = "year") %>% 
  mutate(time = ifelse(year == "first_2020_index",1,ifelse(year == "second_2020_index",2,ifelse(year == "first_2021_index",3,ifelse(year == "second_2021_index",4,ifelse(year == "first_2022_index",5,6))))))

ggplot(cleaned_year_index, aes(x=time, y=value, color = City_name))+
  geom_line(size=1)+
  labs(
    y="index value",
    x="Half year"
  )+
  scale_x_continuous(
    breaks = seq(1,6,1),
    labels = c("2020_1st","2020_2nd","2021_1st","2021_2nd","2022_1st","2022_2nd"),
    limits = c(1,6))+
  scale_color_brewer(palette="Paired")+
  labs(title="Index every half year in according to mean value of first half year of 2020")
```

### Quarterly index


```{r,echo=FALSE,warning=FALSE}
summary_quarter_yr = 
  hp_three_yr %>%
  pivot_longer(col = `43831`:`44927`,names_to = "date") %>%
  mutate(year_quarter = ifelse(date<43912,"first_2020",ifelse(date<44012,"second_2020",ifelse(date<44100,"third_2020",ifelse(date<44190,"fourth_2020",ifelse(date<44280,"first_2021",ifelse(date<44377,"second_2021",ifelse(date<44467,"third_2021",ifelse(date<44557,"fourth_2021", ifelse(date<44647,"first_2022",ifelse(date<44737,"second_2022",ifelse(date<44827,"third_2022","fourth_2022")))))))))))) %>%
  group_by(City_name,year_quarter) %>% 
  summarize(
    mean_val = round(mean(value),2)
  ) %>% 
  ungroup() %>% 
  pivot_wider(
    names_from=year_quarter,
    values_from = mean_val)

quarterly_index = 
  summary_quarter_yr %>% 
  mutate_at(vars(first_2020:third_2022), list(index = ~./first_2020)) %>% 
  select(City_name,ends_with("index")) %>% 
  select(City_name, first_2020_index, second_2020_index, third_2020_index, fourth_2020_index,first_2021_index, second_2021_index, third_2021_index, fourth_2021_index,first_2022_index, second_2022_index, third_2022_index, fourth_2022_index)
quarterly_index[,1:5]
quarterly_index[,6:9]
quarterly_index[,10:13]


```


```{r,echo=FALSE}
cleaned_quarter_index = 
  quarterly_index %>% 
  pivot_longer(col = first_2020_index:fourth_2022_index, names_to = "year") %>% 
  mutate(time = ifelse(year == "first_2020_index",1,ifelse(year == "second_2020_index",2,ifelse(year == "third_2020_index",3,ifelse(year == "fourth_2020_index",4,ifelse(year == "first_2021_index",5,ifelse(year == "second_2021_index",6,ifelse(year == "third_2021_index",7,ifelse(year == "fourth_2021_index",8,ifelse(year == "first_2022_index",9,ifelse(year == "second_2022_index",10,ifelse(year == "third_2022_index",11,12))))))))))))

ggplot(cleaned_quarter_index, aes(x=time, y=value, color = City_name))+
  geom_line(size=1)+
  labs(
    y="index value",
    x="Quarter year"
  )+
  scale_x_continuous(
    breaks = seq(1,12,2),
    labels = c("2020_1st","2020_3rd","2021_1st","2021_3rd","2022_1st","2022_3rd"),
    limits = c(1,12))+
  scale_color_brewer(palette="Paired")+
  labs(title="Index every quarter year in according to mean value of first quarter of 2020")

```


\newpage




## Price distribution for last three month

```{r,echo=FALSE,message=FALSE,warning=FALSE}
time_last = colnames(house_price)[42:54]
east_coast_hp = house_price %>% filter(City_name == "Chappaqua" | City_name == "Scarsdale" |City_name == "Bedford" | City_name == "River Edge")
df_last = t(east_coast_hp[,42:54])
data_df_last = melt(df_last)
data_df_last_named = 
  data_df_last %>% 
  mutate(Var2 = ifelse(Var2==1,"Chappaqua",ifelse(Var2==2, "Bedford", ifelse(Var2==3, "Scarsdale","River Edge"))),
         ind = rep(1:length(time_last),4),
         City_name = Var2)

ggplot(data_df_last_named, aes(x=ind, y=value, color = City_name))+
  geom_line(size=1)+
  labs(
    title="Price distribution for Nov. 2022, Dec. 2022 and Jan. 2023",
    y="House Price",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time_last),3),
    labels = c("Nov_04_2022","Nov_25_2022","Dec_16_2022","Jan_06_2023","Jan_27_2023"),
    limits = c(1,length(time_last)))
```



```{r,echo=FALSE}
all_val_plot = 
  east_coast_hp %>%
  select(-State) %>% 
  select(City_name,"Nov_04_2022":"Jan_27_2023") %>% 
  pivot_longer(
    Nov_04_2022:Jan_27_2023,
    names_to = "date"
  ) %>% 
  mutate(ind = seq(1,length(time_last)*4,1))

ggplot(all_val_plot,aes(x=value, y=ind, color = City_name))+
  geom_point()+
  labs(title = "Scatter plot of house price over all values")
```

Summary of the four cities in the range from November 2022 to January 2023

```{r,echo=FALSE}
all_val_plot %>% 
  group_by(City_name) %>% 
  summarize(
    mean_val = mean(value),
    sd_val = sd(value),
    range_dollar = max(value)-min(value)
  )
```




```

