---
title: "New Report"
output:
  pdf_document: default
  html_document: default
date: "2023-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r,echo=FALSE,message=FALSE}
library("tidyverse")
library("dplyr")
library("readxl")
library(ggplot2)
library(reshape2)
library(ggfortify)
library(wesanderson)
```

```{r,echo=FALSE}
house_price = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 1)
num_sold = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 2)
medium_time = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 3)
all_info = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 4)
pop = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 5)
num_house = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 6)
sale_to_list = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 7)
hp_three_yr = read_excel("neighborhood_comparison/house_price.xlsx", sheet = 8)
```

## Scatter plot for chappaqua and scarsdale

```{r}
all_val = 
  hp_three_yr %>%
  filter(City_name=="Chappaqua" | City_name=="Scarsdale") %>% 
  pivot_longer(
    `43831`:`44927`,
    names_to = "date"
  ) %>% 
  mutate(date_val = as.Date(as.numeric(date),origin="1899-12-30"),
         ind = seq(1,72,1))

ggplot(all_val,aes(x=value, y=ind, color = City_name))+
  geom_point()+
  labs(title="Scatter plot for Chappaqua and Scarsdale")
```


The standard deviation of Chappaqua is ```r round(sd(as.numeric(hp_three_yr[1,][-1:-2])),3)``` with a range of ```r max(as.numeric(hp_three_yr[1,][-1:-2]))-min(as.numeric(hp_three_yr[1,][-1:-2]))``` dollar per square feet. Scarsdale, on the other hand, has a standard deviation of ```r round(sd(as.numeric(hp_three_yr[5,][-1:-2])),3)``` with a range of ```r max(as.numeric(hp_three_yr[5,][-1:-2]))-min(as.numeric(hp_three_yr[5,][-1:-2]))``` dollar per square feet. The table shows all the information:

```{r,echo=FALSE}
all_val %>% 
  group_by(City_name) %>% 
  summarize(
    mean_val = mean(value),
    sd_val = sd(value),
    range_dollar = max(value)-min(value)
  )
```



\newpage



## Mean value for each half year



```{r,echo=FALSE}
summary_half_yr = 
  hp_three_yr %>%
  pivot_longer(col = `43831`:`44927`,names_to = "date") %>%
  mutate(year_half = ifelse(date<44012,"first_2020",ifelse(date<44196,"second_2020",ifelse(date<44377,"first_2021",ifelse(date<44550,"second_2021",ifelse(date<44742,"first_2022","second_2022")))))) %>% 
  group_by(City_name,year_half) %>% 
  summarize(
    mean_val = round(mean(value),2)
  ) %>% 
  ungroup() %>% 
  pivot_wider(
    names_from=year_half,
    values_from = mean_val)
summary_half_yr
```



```{r,echo=FALSE}
cleaned_summ = summary_half_yr %>% 
  pivot_longer(col = first_2020:second_2022, names_to = "year") %>% 
  mutate(time = ifelse(year == "first_2020",1,ifelse(year == "second_2020",2,ifelse(year == "first_2021",3,ifelse(year == "second_2021",4,ifelse(year == "first_2022",5,6))))))

ggplot(cleaned_summ, aes(x=time, y=value, color = City_name))+
  geom_line(size=1)+
  labs(
    y="house price",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,6,1),
    labels = c("2020_1st","2020_2nd","2021_1st","2021_2nd","2022_1st","2022_2nd"),
    limits = c(1,6))+
  scale_color_brewer(palette="Paired")+
  labs(title="Time series plot for mean value every half year")
```


\newpage

## Index

### Yearly index

```{r,echo=FALSE}
yearly_index = 
  summary_half_yr %>% 
  mutate_at(vars(first_2020:second_2022), list(index = ~./first_2020)) %>% 
  select(City_name,ends_with("index")) %>% 
  select(City_name, first_2020_index, second_2020_index,first_2021_index, second_2021_index,first_2022_index, second_2022_index)
  
yearly_index
```

Yearly index plot:

```{r,echo=FALSE}
cleaned_year_index = yearly_index %>% 
  pivot_longer(col = first_2020_index:second_2022_index, names_to = "year") %>% 
  mutate(time = ifelse(year == "first_2020_index",1,ifelse(year == "second_2020_index",2,ifelse(year == "first_2021_index",3,ifelse(year == "second_2021_index",4,ifelse(year == "first_2022_index",5,6))))))

ggplot(cleaned_year_index, aes(x=time, y=value, color = City_name))+
  geom_line(size=1)+
  labs(
    y="index value",
    x="Half year"
  )+
  scale_x_continuous(
    breaks = seq(1,6,1),
    labels = c("2020_1st","2020_2nd","2021_1st","2021_2nd","2022_1st","2022_2nd"),
    limits = c(1,6))+
  scale_color_brewer(palette="Paired")+
  labs(title="Index every half year in according to mean value of first half year of 2020")
```

### Quarterly index


```{r,echo=FALSE,warning=FALSE}
summary_quarter_yr = 
  hp_three_yr %>%
  pivot_longer(col = `43831`:`44927`,names_to = "date") %>%
  mutate(year_quarter = ifelse(date<43912,"first_2020",ifelse(date<44012,"second_2020",ifelse(date<44100,"third_2020",ifelse(date<44190,"fourth_2020",ifelse(date<44280,"first_2021",ifelse(date<44377,"second_2021",ifelse(date<44467,"third_2021",ifelse(date<44557,"fourth_2021", ifelse(date<44647,"first_2022",ifelse(date<44737,"second_2022",ifelse(date<44827,"third_2022","fourth_2022")))))))))))) %>%
  group_by(City_name,year_quarter) %>% 
  summarize(
    mean_val = round(mean(value),2)
  ) %>% 
  ungroup() %>% 
  pivot_wider(
    names_from=year_quarter,
    values_from = mean_val)

quarterly_index = 
  summary_quarter_yr %>% 
  mutate_at(vars(first_2020:third_2022), list(index = ~./first_2020)) %>% 
  select(City_name,ends_with("index")) %>% 
  select(City_name, first_2020_index, second_2020_index, third_2020_index, fourth_2020_index,first_2021_index, second_2021_index, third_2021_index, fourth_2021_index,first_2022_index, second_2022_index, third_2022_index, fourth_2022_index)
quarterly_index[,1:5]
quarterly_index[,6:9]
quarterly_index[,10:13]


```


```{r,echo=FALSE}
cleaned_quarter_index = 
  quarterly_index %>% 
  pivot_longer(col = first_2020_index:fourth_2022_index, names_to = "year") %>% 
  mutate(time = ifelse(year == "first_2020_index",1,ifelse(year == "second_2020_index",2,ifelse(year == "third_2020_index",3,ifelse(year == "fourth_2020_index",4,ifelse(year == "first_2021_index",5,ifelse(year == "second_2021_index",6,ifelse(year == "third_2021_index",7,ifelse(year == "fourth_2021_index",8,ifelse(year == "first_2022_index",9,ifelse(year == "second_2022_index",10,ifelse(year == "third_2022_index",11,12))))))))))))

ggplot(cleaned_quarter_index, aes(x=time, y=value, color = City_name))+
  geom_line(size=1)+
  labs(
    y="index value",
    x="Quarter year"
  )+
  scale_x_continuous(
    breaks = seq(1,12,2),
    labels = c("2020_1st","2020_3rd","2021_1st","2021_3rd","2022_1st","2022_3rd"),
    limits = c(1,12))+
  scale_color_brewer(palette="Paired")+
  labs(title="Index every quarter year in according to mean value of first quarter of 2020")

```


\newpage




## Price distribution for last three month

```{r,echo=FALSE,message=FALSE,warning=FALSE}
time_last = colnames(house_price)[42:54]
east_coast_hp = house_price %>% filter(City_name == "Chappaqua" | City_name == "Scarsdale" |City_name == "Bedford" | City_name == "River Edge")
df_last = t(east_coast_hp[,42:54])
data_df_last = melt(df_last)
data_df_last_named = 
  data_df_last %>% 
  mutate(Var2 = ifelse(Var2==1,"Chappaqua",ifelse(Var2==2, "Bedford", ifelse(Var2==3, "Scarsdale","River Edge"))),
         ind = rep(1:length(time_last),4),
         City_name = Var2)

ggplot(data_df_last_named, aes(x=ind, y=value, color = City_name))+
  geom_line(size=1)+
  labs(
    title="Price distribution for Nov. 2022, Dec. 2022 and Jan. 2023",
    y="House Price",
    x="Date"
  )+
  scale_x_continuous(
    breaks = seq(1,length(time_last),3),
    labels = c("Nov_04_2022","Nov_25_2022","Dec_16_2022","Jan_06_2023","Jan_27_2023"),
    limits = c(1,length(time_last)))
```



```{r,echo=FALSE}
all_val_plot = 
  east_coast_hp %>%
  select(-State) %>% 
  select(City_name,"Nov_04_2022":"Jan_27_2023") %>% 
  pivot_longer(
    Nov_04_2022:Jan_27_2023,
    names_to = "date"
  ) %>% 
  mutate(ind = seq(1,length(time_last)*4,1))

ggplot(all_val_plot,aes(x=value, y=ind, color = City_name))+
  geom_point()+
  labs(title = "Scatter plot of house price over all values")
```

Summary of the four cities in the range from November 2022 to January 2023

```{r,echo=FALSE}
all_val_plot %>% 
  group_by(City_name) %>% 
  summarize(
    mean_val = mean(value),
    sd_val = sd(value),
    range_dollar = max(value)-min(value)
  )
```


